### 2019-05-07 article.tex
Split up in article.tex (big picture), content.tex (the content) and (supplement.tex) supplementary materials

\documentclass{article}

% Target journal: Molecular Phylogenetics and Evolution
%
% From author guidelines:
%
% Short communications of approximately 3000 words are also accepted. 
% These papers should contain no more than two figures, two tables, and thirty references. 
% A short abstract of fewer than 200 words is acceptable.

% Annotation/feedback commands
\newcommand*\rampal[1]{\textcolor{red}{\textbf{[RSE: #1]}}}
\newcommand*\richel[1]{\textcolor{orange}{\textbf{[RJCB: #1]}}}
\newcommand*\gio[1]{\textcolor{green}{\textbf{[GL: #1]}}}

% Bibliography
\usepackage{natbib}
\bibpunct{(}{)}{;}{a}{}{;}

\usepackage{amssymb}
\usepackage[english]{babel}

\usepackage{bbm}
\usepackage{dsfont}

% Use 'It was found that something is something (Name 1234)' style
\setcitestyle{authoryear,open={},close={}}

% Affiliations
\usepackage{authblk}
\title{The error in Bayesian phylogenetic reconstruction when speciation co-occurs}

\author[1]{Giovanni Laudanno}
\author[1]{Rich\`el J.C. Bilderbeek}
\author[1]{Rampal S. Etienne}
\affil[1]{Groningen Institute for Evolutionary Life Sciences, University of Groningen, Groningen, The Netherlands}

% Use double spacing
\usepackage{setspace}
\doublespacing

\usepackage{pgf}
\usepackage{hyperref}
\usepackage{verbatim}
\usepackage{comment}

% Adds numbered lines
\usepackage{lineno}
\linenumbers

\hyphenation{
  BEAST
  Pa-ra-me-ter
  Drum-mond 
  Bayes-ian 
  Mr-Bayes 
  ap-proach-es 
  Rev-Bayes 
  cre-ate
  spe-ci-a-tion-com-ple-tion
  pro-trac-ted
 }

\begin{document}

\maketitle

\include{content}

\include{supplement}

\end{document}