### 2019-06-07 razzo-tables/razzo-table_1.tex

\begin{table}
  \centering 
  \begin{tabular}{r c c c}
    \hline
    Symbol                & Expectation & Expected figure & Measured figure \\
    \hline
    \hline
    $\mathcal{H}_1$       & For a higher $s$, $e$ will increase monotonously & \ref{fig:h_1} & \ref{fig:1} \\
    $\mathcal{H}_2$       & Extinction rates have no effect on $e$ & \ref{fig:h_2} & \ref{fig:2} \\
    $\mathcal{H}_3$       & The timing of multiple-birth events has no effect on $e$ & \ref{fig:h_3} & Not yet \\
    $\mathcal{H}_4$       & An MBD tree will have a higher $e$ than its twin BD counterpart & \ref{fig:h_4} & Not yet \\
    $\mathcal{H}_5$       & The number of taxa has no effect on $e$ & \ref{fig:h_5} & \ref{fig:5} \\
    $\mathcal{H}_6$       & A higher number of taxa decreases the variance in $e$ & \ref{fig:h_6} & \ref{fig:6} \\
    $\mathcal{H}_7$       & The MBD regime has no effect on $e$ & \ref{fig:h_7} & Not yet \\
    \hline
  \end{tabular}
  \caption{
    Overview of the symbols. 
  }
  \label{table:hypotheses}
\end{table}