### 2019-06-19 razzo-discussion.tex
Start of discussion

\section{Discussion}

From the four MBD parameters $\lambda$, $\mu$, $\nu$ and $q$,
we investigated 1, 2, 4 and 3 different values respectively.
We chose to use only one $\lambda$, as the proportion of species
created in a co-occurent speciation event is dependent on the ratio
between $\lambda$ and a combination of $\nu$ and $q$.

We selected our parameters in such a way that the simulated trees 
had usually less than 200 taxa. One could argue that starting from
trees with more taxa would result in a clearer inference, which we
agree upon. We chose to use more replicates over more taxa, 
as we could easily add more replicates in a scheduled way.