### 2019-07-01 razzo-results.tex

\section{Results}

The inference error made for each of the parameter combinations
is shown in Fig.~\ref{fig:results}. For both extinction rates,
we find that the error increases as $\nu$
and $q$ increase. Also in line with our predictions, we find no difference
between the two extinction rates \giovanni{I don't actually have predictions on that.}.

\input{razzo-figures/fig_results.tex}