### 2019-07-01 razzo-discussion.tex

\section{Discussion}

From the four MBD parameters $\lambda$, $\mu$, $\nu$ and $q$,
we investigated 1, 2, 4 and 3 different values respectively.
We chose to use only one $\lambda$, as the proportion of species
created in a co-occurent speciation event is dependent on the ratio
between $\lambda$ and a combination of $\nu$ and $q$.