### 2019-07-15 article.tex
'speciation completion' are two words

\documentclass{article}
\usepackage{natbib} % Bibliography
\bibpunct{(}{)}{;}{a}{}{;} % Bibliography
\setcitestyle{authoryear,open={},close={}} % Use 'It was found that something is something (Name 1234)' style

% Comments
\newcommand*\rampal[1]{\textcolor{red}{\textbf{[RSE: #1]}}}
\newcommand*\richel[1]{\textcolor{orange}{\textbf{[RJCB: #1]}}}
\newcommand*\giovanni[1]{\textcolor{blue}{\textbf{[GL: #1]}}}

% Packages
\input{razzo-packages.tex}

\hyphenation{
  BEAST
  Pa-ra-me-ter
  Drum-mond 
  Bayes-ian 
  Mr-Bayes 
  ap-proach-es 
  Rev-Bayes 
  cre-ate
  spe-ci-a-tion
  com-ple-tion
  pro-trac-ted
 }

\input{razzo-header.tex}

\begin{document}

\maketitle

\input{razzo-abstract.tex}

\input{razzo-introduction.tex}

\input{razzo-methods.tex}

\input{tab_simparameters.tex}

\input{razzo-results.tex}

\input{fig_results.tex}

\input{razzo-discussion.tex}

\bigskip\bigskip

\bibliographystyle{mee}
\bibliography{razzo-bibliography.bib}

\bigskip\bigskip

\input{razzo-supplement.tex}

\end{document}