### 2019-10-07 razzo-results.tex
Depend on master branch

\section{Results}

The inference error made for each of the parameter combinations
is shown in Figures ~\ref{fig:errors_yule} and ~\ref{fig:errors_bd}. 
For both extinction rates,
we find that the error increases as $\nu$
and $q$ increase. 
Also in line with RJCB's predictions, 
we find no difference between the two extinction rates.