### 2019-11-27 razzo-results.tex

\section{Results}

The inference error made for each of the parameter combinations
is shown in Figures ~\ref{fig:errors_yule} and ~\ref{fig:errors_bd}.

\richel{if true:}The lowest errors are at settings where $nu = 0 \vee q = 0$,
confirming $\mathcal{H}_1$.

\richel{if true:}For both extinction rates,
we find that the error increases as $\nu$
and $q$ increase, confirming $\mathcal{H}_2$

\richel{if true:}There is no difference between the
errors for the two different extinction rates, which
confirms $\mathcal{H}_4$

[other hypotheses being rejected/accepted here]