### 2020-03-31 fig_results.tex
Use nicer figures

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{20200126_figure_1a.png}
  \caption{
    The error distribution for a extinction rate $\mu = 0.0$.
  }
  \label{fig:errors_yule}
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{20200126_figure_1b.png}
  \caption{
    The error distribution for extinction rate $\mu = 0.1$.
  }
  \label{fig:errors_bd}
\end{figure}