### 2020-03-31 fig_results.tex
Moved all figures to main text

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{20200204_fig_esses.png}
  \caption{
    Effective sample sizes of the likelihood of all the experiments that
    finished within 10 days. 
    Each point is one parameter setting.
  }
  \label{fig:esses}
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{20200204_fig_n_taxa.png}
  \caption{
    Histogram of the number of taxa in the full experiment.
  }
  \label{fig:n_taxa}
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{20200204_figure_1a.png}
  \caption{
    The error distribution for a extinction rate $\mu = 0.0$.
  }
  \label{fig:errors_yule}
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{20200204_figure_1b.png}
  \caption{
    The error distribution for extinction rate $\mu = 0.1$.
  }
  \label{fig:errors_bd}
\end{figure}