### 2020-03-31 razzo-results.tex

\section{Results}

The inference error made for each of the parameter combinations
is shown in Figures ~\ref{fig:errors_yule} and ~\ref{fig:errors_bd}.
The two figures, that only differ in their extinction rate, 
show the same error distributions. There is no clear difference between
the true and twin error distribution. For $\nu \leqslant 0.5$, the
range of errors, however, differ. For $\nu \geq 1.0$, the error
distributions become bimodal.

All intermediate data can be downloaded from \url{https://richelbilderbeek.nl/razzo_project_20200204.zip}.