### 2020-03-31 razzo-results.tex

\section{Results}

98% of all runs finished within 10 days. 
Of these runs, figure \ref{fig:esses} shows the distribution 
of their effective samples sizes, with 88\% above the recommended 
value of 200.

Figure \ref{fig:n_taxa} shows the number of taxa for all runs. 
Most runs have less than 100 taxa, of which the median and mean both lie
beloe 30 taxa.

The inference error made for each of the parameter combinations
is shown in Figures ~\ref{fig:errors_yule} and ~\ref{fig:errors_bd}.
The two figures, that only differ in their extinction rate, 
show the same error distributions. There is no clear difference between
the true and twin error distribution. For $\nu \leqslant 0.5$, the
range of errors, however, differ. For $\nu \geq 1.0$, the error
distributions become bimodal.

All simulated data can be downloaded from \url{https://richelbilderbeek.nl/razzo_project_20200204.zip}.