### 2020-04-03 fig_results.tex
Changes @rsetienne

\begin{figure}[!htbp]
  \includegraphics[width=1.0\textwidth]{20200204_fig_esses.png}
  \caption{
    Frequency distribution of effective sample sizes of all experiments that
    finished within 10 days. 
    Each point represents one parameter setting.
  }
  \label{fig:esses}
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=1.0\textwidth]{20200204_fig_n_taxa.png}
  \caption{
    Frequency distribution of the number of taxa across all experiments.
  }
  \label{fig:n_taxa}
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{20200204_figure_1a.png}
  \caption{
    The inference error distribution for a generating extinction rate of $\mu = 0$.
  }
  \label{fig:errors_yule}
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{20200204_figure_1b.png}
  \caption{
    The inference error distribution for a generating extinction rate of $\mu = 0.1$.
  }
  \label{fig:errors_bd}
\end{figure}