### 2020-04-03 razzo-results.tex

\section{Results}

Of all runs 98\% finished within 10 days. Of these runs, 88\% had effective sample sizes above the recommended value of 200, and most well above this (Fig. \ref{fig:esses}). Most runs had less than 100 taxa, of which the median and mean both lie below 30 taxa (Fig. \ref{fig:n_taxa}).

We find no clear difference between the error distributions for the true and twin trees, regardless of whether there is extinction or not in the generating process (Fig.~\ref{fig:errors_yule} and Fig. ~\ref{fig:errors_bd}. We do observe that for generating values of $\nu \leqslant 0.5$, there is only one mode close to 0, whereas for $\nu \geq 1.0$, the error
distributions become bimodal, with a second mode around 0.25. 
All simulated data can be downloaded from 
\begin{sloppypar}
\url{https://richelbilderbeek.nl/razzo_project_20200204.zip}.
\end{sloppypar}